---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# densdpqr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of densdpqr is to make a minimal package for applying the`[dpqr]*()` distribution functions (e.g. `dnorm()`, `pnorm()`, `qnorm()`, `rnorm()`), for any arbitrary output from the `stats::density()` function.

## Example

1. Call `stats::density()` on your data to get an estimate for a density function.
2. Call `densdpr::smooth_density()` to smooth your density function and create the needed distribution functions for `ddens()`, `pdens()`, `qdens()` and `rdens()`.
3. Pass output from `smooth_density()` and other required parameter into `[dpqr]dens()` function(s).

```{r example}
# devtools::install_github("brshallo/densdpqr")
library(densdpqr)

set.seed(123)
data <- rnorm(10000)
dens <- density(data)
sdens <- smooth_density(dens)

ddens(sdens, 2)
pdens(sdens, 2)
qdens(sdens, 0.977)
rdens(sdens, 10L)
```

# Problems

* `smooth_density()` uses a separate smoothing spline to build the requisite functions for `ddens()`, `pdens()` and `qdens()`. This is inefficient *and* means it is possible for the different distribution functions to be slightly inconsistent with one another. What likely *should* be done is... 
    * `ddens()` should serve as the foundation. 
    * `pdens()` would be the integral of `ddens()`. 
    * `qdens()` would be the inverse of `pdens()`. 
    * This would ensure that all distribution functions are consistent with one another^[When I tried to pass `lower = -Inf` into `integrate()` I got errors that the function was divergent... so would need to change the smoothing method in some way to prevent this. I also wasn't sure how to efficiently return the inverse of `pdens()`.].
* May want to change fit of smoother to one that is [orthogonal](https://en.wikipedia.org/wiki/Cubic_Hermite_spline)^[Current method may be biased towards smoothing density rather than the inverse, the quantiles.]
* `qdens()` should not accept values outside of 0 and 1, but it currently does; `pdens()` should not return values outside of 0 to 1, but it can.
    * This could just be hard capped, but it would be preferable to use a smoothing function that respected these density constraints more intelligently^[In the interim, maybe should at least at a `warning()` and just count on user being thoughtful.].
* The smoothing method in `smooth_density()` will not necessarily respect the `from` and `to` ranges set when creating `density()`^[Again are just relying on user to be intelligent when using.].
* Not tested on wide range of distributions.

# Inspiration & Other Resources

Thank you to [spatstat](https://github.com/spatstat/spatstat.core/blob/76d20a642867d7c30a0f11c58af5b8634ad302cd/R/quantiledensity.R) authors as well as William Huber whose comment on [How to find probability density function from density function...](https://stats.stackexchange.com/a/553271/193123) inspired the approach in densdpqr.

* [logspline](https://cran.r-project.org/web/packages/logspline/logspline.pdf) package
* [ks](https://cran.r-project.org/web/packages/ks/ks.pdf) package
* [Nonparametric Statistics, Kernel density estimation...](https://bookdown.org/egarpor/NP-UC3M/kde-i.html) online book

In most simple univariate cases I would just use the `logspline::[dpqr]logspline()` functions which are set-up the same way as `densdpqr::[dpqr]dens()` but without the same [Problems]. The advantages with densdpqr are it allows the use of the base `stats::density()` function, it has fewer dependencies, and is [likely faster](https://gist.github.com/brshallo/ea2e04347e14fae7ff969a54e2266359).
  
  


